Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-28 Response.tex

-To produce constant thrust magnitude or no thrust magnitude (typical of nuclear electric propulsion, or very simple electric propulsion systems.)

-To produce variable thrust magnitude associated with various throttle points at different power accessibility levels (solar electric propulsion).
That is not to say that useful solutions cannot be extracted from the problem, as the authors have shown.  However, a great many solutions in this approach will be unusable due to the fact that acceleration and thrust magnitude are different, i.e. a=T/m

Thus - a MAJOR focus of future work should be the introduction of a seventh state (mass).  This gives rise to a more challenging optimal control problem requiring the switching function, but it is the form that is generally needed in computing optimal open-loop trajectories that can actually be flown.
\end{itshape}

Thank you for the comment and more focus has been included on the pros and cons of this proposed method.
Additional detail has been added to the conclusion which discusses the advantages and the disadvantages, as well as the avenues for future research.
The conclusion is copied below.
\begin{correction}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

@@

\item 
    \begin{itshape}
The pros and cons of the PCRTBP

-The PCRTB gives rise to exploiting Poincare sections in a systematic manner.  2D solutions might be close enough for qualitative understanding of the solution space for many problems (as long as large plane changes aren't needed).

-The PCRTB is limiting for many pragmatic departure orbits (e.g. LEO and GTO, the most common departure orbits for low-thrust transfers).
\end{itshape}

Thank you for the comment and these characteristics of our choice of the PCRTBP. The authors would like to point out that the key idea of designing low-thrust orbits through the reachability analysis on the \Poincare section may be applied to other dynamic system. This manuscript illustrates such idea through PCRTB.

However, according to the reviewers suggestion, additional details have been stated in the problem formulation section, including discussions on the choice of the PCRTBP and the relative strengths/weaknesses of this approach as copied below.

@@

\begin{enumerate}
    \item
        \begin{itshape}
            In the statement "In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential
            missions at a reduced energy cost [15]" - you may want to reference, for example, the Lunar Ice Cube mission paper which absolutely requires low-thrust propulsion and non-keplerian dynamics to satisfy its mission objectives.

            Folta, D., Dichmann, D., Clark, P., Haapala, A., and Howell, K., "LunarCube Transfer Trajectory Options," AAS/AIAA Space Flight Mechanics Meeting, Williamsburg, Virginia, January 11-15, 2015.
        \end{itshape}

        Thank you for the comment and the additional reference.
        In addition to this change we have made other modifications to the introductory paragraphs to improve the organization and clarity.
        Specifically, the first paragraph has been divided, which motivates the use of electric propulsion and the advantages of small satellites which use electric propulsion.
        In addition, the suggested reference has been included in the paper.
        The modified paragraphs are duplicated below.

        \begin{correction}
            Designing spacecraft trajectories is a classic and ongoing topic of research.
            There has been significant research into the design of orbital transfers for space vehicles.
            Optimal expenditure of onboard propellant is critical to allowing a mission to continue for a longer period of time or to enable the launch of a less massive spacecraft.
            Electric propulsion systems offer a much greater specific impulse than chemical systems.
            As a result, the greatly increased efficiency allows for greater payload mass or extended duration missions.
            However, these electric propulsion systems typically have much less thrust than their chemical counterparts and therefore orbital maneuvers have a much longer time of flight.
            In spite of this drawback, a wide variety of missions, such as communication and deep space probes, have utilized the unique benefits of low thrust electric propulsion to great effect~\cite{choueiri2009}.

            With reduced development intervals and decreased launch costs, small satellites have gained increased attention as a cost effective means of scientific and technologic development. 
            Furthermore, these cubesats are easily launched as secondary payloads rather than requiring a dedicated launch.
            The merger of small satellites with miniature electric propulsion enables inexpensive and responsive missions requiring large changes in orbital energy or extended mission lifespan.
            Recent developments in miniature electric propulsion offer the potential for new research opportunities for small spacecraft~\cite{haque2013}.
            The advancements in miniature electric propulsion allows for greater flexibility in the deployment of cubesats as secondary payloads.
            Without the explicit requirement of a dedicated launch trajectory, cubesats are able to exploit a much wider range of possible mission designs~\cite{folta2015}.
            With the potential for more demanding missions, even greater importance is placed on the mission design to ensure that optimal trajectories satisfy mission requirements. 
            In addition, non-Keplerian orbits and multi-body dynamics have been shown to allow for a much greater range of potential missions at a reduced energy cost~\cite{folta2015,koon2011}.
            Future space missions are increasing in complexity and will require new classes of orbits that are not possible via the traditional patched-conic approach~\cite{ross2006,gomez2001}.
            Optimally combining the structure of the dynamic environment with low-thrust propulsion systems is vital for future mission success.
        \end{correction}

@@

\begin{correction}
Finally, there is no closed form analytical solution for the three-body problem and there are an insufficient number of analytical constants~\cite{szebehely1967}.
As a result numerical methods are often required to investigate solutions to the three body problem.
\end{correction}

\item 
    \begin{itshape}
"These techniques suffer from numerical instability and energy drift behaviors which make them ill-suited for long-term propagation. These dissipative effects are even more detrimental with the addition of low-thrust propulsion to the dynamic equations of motion."  It really depends on what you mean by "long term".  If transfers are on the order of hundreds of days, then high precision runge-kutte integrator, for example a Runge-Kutte Verner 8(9), can maintain suitable accuracy such that the energy drift in the Hamiltonian is very small, as can be seen in some of the references that the authors have cited.  If, on the other hand you mean 10-100's of years or more, i.e., when you create your Poincare maps, then this drift is more apparent.  I think that this statement needs to be quantified better because it gives a dishonest portrayal of high-precision Runge-Kutte integrators without numbers attached to it.
\end{itshape}

Thank you for the comment. Those numerical dissipations would depend on several parameters. But, for the particular numerical comparison illustrated by \Cref{fig:compare_energy}, such difference in conservation properties becomes noticeable in the simulation for 15 years. As suggested, the claims about the variational integrator have been qualified with some additional detail.
The correction is copied below.
\begin{correction}
As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
\end{correction}

\item
    \begin{itshape}
"Conventional integration techniques fail to capture the physical laws and geometric properties of the dynamic system [8]. As a result, the long term effects of low-thrust on the spacecraft trajectory are not accurately captured."  Again, I think situationally I would disagree.  Transfers on the order of 100's of days can maintain accuracies of 1e-10 (non-dimensional units) in the Hamiltonian, which is generally considered a "small" amount of energy drift.
\end{itshape}

These sentences have been modified and combined to enhance clarity and reduce some of the unnecessary bloat. 
In addition, some additional detail has been added to explicitly define the notion of ``long-term'' in the context of numerical integration. As shown by \Cref{fig:compare_energy}, the variational integrator exhibit smaller conservation errors than ODE45 at the similar computational efforts, for the particular numerical case considered. The modifications are copied below.
\begin{correction}
    Conventional Runge-Kutta integration techniques, such as those implemented in~\cite{mingotti2011,grebow2011}, fail to capture the physical laws and geometric properties of the dynamic system.
    As a result, these methods suffer from numerical instability and energy drift behaviors which make them ill-suited for the long-term propagation, e.g. on the order of decades to centuries, which is typical of the three-body problem. 
\end{correction}

@@

Thank you for the comment.
The PCRTBP was primarily chosen as a simple dynamic system which still captures the primary structure of the three-body problem.
The introduction has been modified with an additional sentence indicating our use of the planar three body problem

\begin{correction}
Utilizing the planar assumption greatly simplifies the problem and allows for the use of established methods in determining periodic orbits and \Poincare sections.
\end{correction}

\item 
    \begin{itshape}
-"In addition, through the use of geometric integrators we accurately capture the effects of low thrust on the system dynamics in the numerical simulation."  A variational integrator will accurately capture the qualitative effects of the transfer (e.g. propellant use), however it will not necessarily accurately capture accuracy in the final state vector - it is guaranteed to be on an energy surface, but it is not guaranteed to be at exactly the right location on the energy surface.  I would soften this claim or reword.
\end{itshape}

While the geometric integrator will not exactly characterize the terminal energy or state of a dynamic system, it will better preserve the total energy variation as compared to conventional integration schemes. 
In addition, the total energy deviation will remain bounded and less than than of conventional schemes which experience unbounded energy deviation. 
As a result, the terminal state/energy will be more accurately characterized using a variational integrator as compared to conventional schemes. 
Also this difference becomes more evident over increasing time horizons.
Some additional detail has been added stating that the increased accuracy is most evident over long time periods.

@@

\begin{correction}
In addition, through the use of geometric integrators we more accurately capture the effects of low thrust on the system dynamics in the numerical simulation over extended periods which are typical of electric propulsion systems. 
\end{correction}

\item 
    \begin{itshape}
-"The Poincare section reduces the dimensionality of the system dynamics to the study of a related discrete update map." It does, but at the expense if a large up-front computational run that must be run well in advance to capture the problem characteristics.  Also, it is more challenging to execute in 3D, which should be noted.
\end{itshape}

@@

Thank you for the comment.
Some additional detail has been included mentioning the difficulties in applying the \Poincare~section approach to the 3D case.

\begin{correction}
While the \Poincare~section is well defined for the planar case, as a 2D representation, it is more challenging to implement and visualize for the general 3D three-body problem.
\end{correction}

\item
    \begin{itshape}
-"This discrete update map, or variational integrator, shares the same geometric properties of the continuous time system and exhibits much better energy behavior than the traditional integration methods, especially over long transfer durations with small magnitude control inputs."  Yes, but the order of the variational integrator still matters.  The drift in the state vector accuracy is still important to a mission designer, even if it is guaranteed to lie on an energy surface.  I think you should indicate that this is more for preliminary analysis to characterize the transfers, when this drift in state vector accuracy is tolerable.  I.E., it's useful for the introductory phases of mission design when trying to globally characterize the design space.  Characterizing the scope of the analysis in this way would be helpful to the audience.
\end{itshape}

Thank you for the comment.
The section on the variational integrator has been reduced and this specific sentence has been removed. 
However, additional detail has been provided in the variational integrator subsection which discusses this drawback.

@@

\end{correction}
\item 
    \begin{itshape}
"Optimal solutions are generally sensitive to small variations in the initial multipliers. As a result, the numerical stability of sensitivity derivatives is critical to accuracy and computational performance."  I don't agree that this implies that variational integrators are better for gradient-based methods.  The primary issue is that gradient-based methods that are solving optimal control problems are using first-order approximations to iterate on the problem.  The first-order approximation is what is poor, not necessarily the gradients.  Even if these gradients are of exact accuracy (and I agree that they should always be as accurate as possible), the first-order methods in the chaotic problem are what give the difficulties.  One could use a variational or standard integrator in a first-order gradient based method and nothing will change the fact that the method will be highly sensitive, save for partitioning the sensitivities as you did with your multiple shooting approach.  At best, you can change the statement to indicate that the accuracy in the gradients may improve, which is always helpful to gradient-based methods, but it's no guarantee that it eliminates major problem sensitivity.
\end{itshape}

It has been shown that the structure-preserving properties of geometric numerical integrators are also useful to numerical optimizations for improved computational efficiency and stability. For example, see the following reference.

\begin{itemize}
\item[-] Ober-Blöbaum, Sina, Oliver Junge, and Jerrold E. Marsden. "Discrete mechanics and optimal control: an analysis." ESAIM: Control, Optimisation and Calculus of Variations 17.2 (2011): 322-352.
\end{itemize}

The proposed approach utilizes desirable numerical properties of the variational integrator. However, the main contribution is systematic design of low-thrust orbital maneuvers through reachability set analysis on the \Poincare section. The above sentences may be confusing, and the authors decided to remove them from the introduction of the revised manuscript.

@@

Thank you for the comment.
All of the figures showing position states have been modified to use \texttt{axis equal}.

\item 
    \begin{itshape}
Section 3.1 has been discussed in many references, the reader may want to direct them to this material.
(Choose one of Marsden or West's papers for example.)
\end{itshape}

Thank you for the comment.
The discussion on the variational integrators has been reduced. 
Much of the background material on Lagrangian mechanics is removed and an additional paragraph is included to summarize the key ideas of variational integrators. 
In addition, the variational integrator details are moved into a titled subsection.
The additional paragraph is shown below.

\begin{correction}
Geometric numerical integration deals with numerical integration methods which preserve the geometric properties of the flow of a differential equation, such as invariant properties and symplecticity.
Variational integrators are constructed by discretizing Hamilton's principle rather than the continuous Euler-Lagrange equations.
As a result, integrators developed in this manner have the desirable properties that they are symplectic and momentum preserving.
In addition, they exhibit improved energy behavior over long integration periods.
A thorough discussion of variational integrators is provided in~\cite{west2004,marsden2001}.
We use this approach to construct a variational integrator for the PCRTBP with the inclusion of low-thrust propulsion.

@@

We chose the trapezoidal rule as it provides a second order quadrature approximation.
Furthermore, instead of using Taylor methods or higher order options, the trapezoidal rule provides an explicit second order accurate integration rule.
This greatly reduces the complexity of the optimization routine.
In addition, the background material on the variational integrators have been removed from the manuscript. 
Only the specific details covering the development of the variational integrator for the PCRTBP are discussed.

\item 
    \begin{itshape}
Section 3.2 - please cite a reference as to where the selected quadrature rules presented in Table 1 have been derived.
\end{itshape}

In the interest of conciseness and to avoid repeating material that has already been covered in the literature, the discussion on variational integrators has been substantially reduced.
The background material on the variational integrators has been removed and this table has also been removed from the manuscript.

\item 
    \begin{itshape}
Figure 2 - x-axis has no label but I think it's non-dimensional time, please indicate.  Label all axes even if they are non-dimensional units.
\end{itshape}

@@

Thank you for the comment. 
All of the plots have been updated to include all axes labels.
Furthermore, the specific plot in questions has been replaced with another plot more explicitly showing the benefits of variational integrators over conventional integration schemes.
This modified plot is shown below.

\begin{correction}
    \begin{figure}[H] 
        \centering 
        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparison: A stable trajectory is used to test the variational and conventional integrators. 
        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
        [0.5\textwidth]{\includegraphics[width=\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
\end{figure}
\end{correction}
\item 
    \begin{itshape}
I recommend removing appendix B.  Giving the equations and stating that the Matrix inversion is from Gauss-Jordan elimination is probably sufficient, unless there is some compelling algebraic manipulation that is very unique to the problem.
\end{itshape}

@@

For example -

Figure 2.4
Lew, A., Marsden, J., Ortiz, M., West, M., "An Overview of Variational Integrators," FINITE ELEMENT METHODS: 1970's AND BEYOND L.P. Franca (Ed.) c CIMNE, Barcelona, Spain 2003.

Online copy:
http://lagrange.mechse.illinois.edu/pubs/LeMaOrWe2004/LeMaOrWe2004.pdf
\end{itshape}

### 2017-07-28 Response.tex

Thank you for the comment.
The comparison plot has been modified to follow the precedent in the literature.
The plot has been modified to show the energy deviation as a function of computational time for both the variational integrator and \texttt{ODE45}.
The modified content is copied below.
\begin{correction}
\begin{figure} 
        \centering 
        \begin{subfigure}[h]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{trajectory} 
                \caption{Earth-Moon three-body trajectory used for integrator comparison}
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
%       \begin{subfigure}[htbp]{0.3\textwidth} 
%               \includegraphics[width=\textwidth]{components} 
%               \caption{State \( \parenth{x , y, \dot{x}, \dot{y}}\)} \label{fig:compare_components} 
%       \end{subfigure} ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=\textwidth]{figures/cputimevsE.pdf} 
                \caption{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}} 
        \end{subfigure} 
        \caption{Integrator Comparison}
\end{figure}
A simulation comparing the variational integrator to a conventional Runge-Kutta method is given in~\cref{fig:integrator_compare}.
A particle is simulated from an initial condition of \( \vecbf{x}_0 = \begin{bmatrix} 0.75 & 0 & 0.2 & 0\end{bmatrix}^T \) for \( t_f = 200 \approx 15\) years in the Earth-Moon system.
    The variational integrator uses a range of step sizes between \SIrange{47.2}{4720}{\second} while the Runge-Kutta method uses a variable step size implemented via \texttt{ODE45} in Matlab.
    The step size of the variable integrator is varied to approximately match the run time required by the conventional \texttt{ODE45} integrator.
\Cref{fig:compare_trajectory} shows the trajectory of the spacecraft in the rotating reference for this comparison.
Both integration schemes result in trajectories that are initially nearly identical.
\Cref{fig:compare_energy} shows the mean Jacobi integral deviation over the entire simulation time as a function of computation time.
For a given computational effort, in the form of simulation run time, the variational integrator will provide a smaller energy deviation as compared to the conventional integration scheme.
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
\end{correction}

@@

\item 
    \begin{itshape}
It may also be generally worth acknowledging the benefits of Taylor's method within the context of variational integrators:
Schmitt, J., Shingel, T., Leok, M., "LAGRANGIAN AND HAMILTONIAN TAYLOR VARIATIONAL INTEGRATORS," eprint arXiv:1703.06599, 2017

Online copy: https://arxiv.org/pdf/1703.06599.pdf 
\end{itshape}

Thank you for the comment.
However, this manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
There is a wide range of topic under the umbrella of variational integrators that are not directly applicable to this manuscript and work.
Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.
As a result, we do not feel it is applicable to include additional discussion on the merits of taylor integration methods.

@@

Thank you for the comment.
The additional clarification has been added to the manuscript and copied below.
\begin{correction}
Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
\end{correction}

\item 
    \begin{itshape}
Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.  This needs to somehow be conveyed.  3D transfers are significantly more challenging when it comes to exploiting Poincare sections.  It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.
\end{itshape}

Thank you for the comment.
An additional paragraph has been added to the beginning of the mathematical background section which discusses our reachability based transfer approach.
This additional content is copied below.

\begin{correction}
The numerical examples presented in this section are designed in the context of the PCRTBP.
The dynamic environment has a four dimensional state space and offers a convenient integration constant in the form of the Jacobi integral.
As a result, there are well defined methods to define and exploit \Poincare sections, which result in straightforward two-dimensional subspaces of the system.
Our approach uses the \Poincare section to approximate the reachability set on this reduced subspace.
As a result, this approach is more difficult to apply to three-dimensional transfers in the general three body problem.
\Poincare sections in the case of the general six dimensional state space are significantly more challenging and typically require more complicated visualization techniques. 
However, this is an area of active research and some of the authors future research is aimed at implementing this approach for non-planar transfer trajectories~\cite{kulumani2016d}.
\end{correction}

@@

\item
    \begin{itshape}
Figures 10D and 13D - the authors need to present these plots in dimensional coordinates so that they may be reconciled with what existing technology can produce.  I recommend Newtons or milliNewtons.  For example, the NEXT (NASA evolutionary xenon thruster) can produce 0.25 Newtons of thrust at maximum power (~7 kW).  The XR5 Hall thruster can produce comparable thrust at 4.5 kW of power.  These are currently the highest-power single engines other than the ARM Hall thruster which is not flight ready yet.  For these examples to be reconciled against the capability of existing technology, we need to see the required forces that are being demanded from the optimal control problem.  It may be that the results are beyond the scope of current technology - that's fine - just say that they are examples and that it's ongoing work to formulate example problems that capture existing technology better.
\end{itshape}

Thank you for the comment.
The mentioned plots of the control input have been modified to show the data in dimensional units (Newtons). 
Some additional detail has been provided in the manuscript which discusses the choice of \( u = 0.75 \) and it's relationship to the current state of the art thruster systems.
The modified plots are not included here but the additional content is copied below in response to a different but related comment.

\begin{correction}
Next, we determine the reachability set with the addition of a low thrust control input.
We define a maximum magnitude of the thrust as \( u_{max} = 0.75 \) and assume we can point the thrust in any direction within the plane. 
This model is representative of many spacecraft which have a body fixed thruster and attitude control system.
We assume a fully actuated spacecraft model which decouples the translational and rotational dynamics.
This acceleration limit is approximately \( u_{max} = \SI{2}{\milli\meter\per\second\squared} \) in the Earth-Moon system.
Assuming a fixed spacecraft mass of \SI{500}{\kilo\gram}, this model defines a maximum thrust of \SI{1}{\newton}.
Currently, the NASA NEXT xenon thruster is able to provide approximately \SI{0.25}{\newton} of thrust, and a cluster of such engines could be used to provide the desired thrust used in this work~\cite{schmidt2008}.
The trajectories are generated from a fixed initial state of \( \vecbf{x}_0 = \begin{bmatrix}0.8156 & 0 & 0 & 0.1922 \end{bmatrix}^T \) over a fixed time span of \( t_f = 1.4 \).
This initial state lies on the initial periodic orbit and the time of flight is equivalent to one half period of the initial periodic orbit. 
\begin{figure}[H]
    \centering
    \subcaptionbox{Control input for the selected transfer trajectory}{\includegraphics[width=0.5\textwidth]{control_input_l1}}~
    \subcaptionbox{Control Input: Combined control history for each stage of the reachability analysis. 
    The control always remains within the maximum bound.}{\includegraphics[width=0.5\textwidth]{figures/control_input_geo.pdf}}
\end{figure}
\end{correction}

@@

Thank you for the comment.
Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
The modified portion is copied below.

\begin{correction}
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
\end{correction}

\item
    \begin{itshape}
Did the authors experiment with the number of multiple shooting stages in their example problems?  There must be some sort of "sweet spot" in number of stages for the example problems attempted.  If I missed it, I apologize, but I didn't see it.
\end{itshape}

We did spend considerable time in determining the appropriate number of stages. 
One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
For simplicity, this reduced the possible number of stages to an even number. 
Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in convergence. 
The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.

@@

\item 
    \begin{itshape}
Is this method exceptionally difficult to adapt to create a numerical example from LEO or GTO?  These would be the most pragmatically interesting departure orbits to the mission design community.  Section 6.2 presents an example from GEO, but it is unlikely that GEO represents a "typical" parking orbit, since there is no reason to insert a spacecraft into GEO before it transfers to the destination orbit.  This should be called out in the conclusions or future work.
\end{itshape}

This method is not difficult to adapt for other scenarios.
However, there is a certain amount of experimentation to ensure that trajectories converge. 
In addition, the details in this paper were designed in the context of the planar three body problem and as a result and extension to the nonplanar GTO is non trivial.
The examples presented are primarily aimed at introducing and demonstrating the feasibility of this approach, rather than designing operational mission profiles for spacecraft.
However, this is a valuable comment and a goal of future research. 
Additional detail has been included in the conclusion discussing this topic.
The modified conclusion is copied below.

\begin{correction}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare~section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare~section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
Direct optimal control techniques must rely entirely on the ability of the numerical optimization routine to determine a feasible solution.
Our approach leverages the benefits of the reachability set, which encompass the maximum set of states achievable by the spacecraft over a fixed time horizon.
Using the reachability set on a \Poincare section reduces the dimensionality of the system and simplifies the analysis by avoiding the cost of a completely unmotivated exploration of the space.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

@@

There is additional research to extend these results to more general transfer scenarios in future work.
The incorporation of fourth body perturbations, such as the Sun in the Earth-Moon system, offers an additional method of increasing the reachable set with the combined use of the solar perturbation and low-thrust propulsion.
In addition, the assumed acceleration magnitude is currently beyond the capabilities of current electric propulsion systems.
Furthermore, this analysis did not consider the effect of variable mass on the optimal control solution.
This will result in a more complicated optimal control problem and is a focus of future research.
Finally, Lyapunov control theory, which has previously been applied to the two-body problem, is being investigated in the hope of designing closed loop control schemes for this three-body scenario~\cite{chang2002}.
The addition of attitude dynamics and realistic pointing constraints would significantly improve the applicability of this work.
\end{correction}

\item 
    \begin{itshape}
The authors should discuss some statistics of the convergence of their multiple shooting TPBVP - number of iterations, compute time, etc.
\end{itshape}

Thank you for the comment, however it seems to be rare for the research community to give detailed description of the numerical optimization method.
Most of the literature tends to not spend time discussing the low level details of the computational procedure, e.g.~\cite{grebow2011,stuart2010}.
In addition, the number of iterations, computation time and even the eventual cost function value are highly dependent on the specific system and more specifically the implementation of the algorithm in software. 
There are a wide variety of works which are focused more exclusively on the computational implementation of various optimal control techniques which do spend significant time on the computational characteristics.
However, our manuscript is primarily focused on demonstrating the feasibility of using reachability sets to design transfer trajectories.
As a result, there is not much benefit to detailed statistics of the implementation of this methodology, and we have omitted it from the manuscript.
However, we have added some rough detail on the computation time and computer hardware used in the numerical optimization.
This additional detail is copied below.

@@

\begin{correction}
Each optimal control solution, corresponding to a discrete value of \( \theta_d \), is solved using \texttt{fsolve} as described earlier. 
Each solution on the reachability set is computed in approximately \SI{30}{\second} on a desktop computer using an \SI{3.4}{\giga\hertz} Intel i7-3700.
\end{correction}

\item 
    \begin{itshape}
Can any intuition be gained from reporting the optimal converged cost function values in section 6.1 and 6.2?  One could at least compare them to compare the reachability in the two different examples.
\end{itshape}

The cost function only parameterizes the distance between the controlled and uncontrolled terminal states on the \Poincare section. 
In addition, each state which lies on the reachable set would typically have a different distance and therefore a different converged cost function value. 
Furthermore, it would be difficult to characterize a single parameter which characterizes the reachable set, in the form of the cost function, for all states of the reachable set. 
Also, the cost function is highly dependent on the transit time of trajectory, i.e. a longer transit time corresponds to a larger reachable set and a larger magnitude of the cost function.
However, this is an interesting suggestions for future research.
For example, the cost function gives some indication of area/volume of the reachable set. 
It may be possible to use this value as another way to characterize the subspace on the \Poincare section and is an intriguing avenue of future research.

@@

-Presented control magnitudes are dimensionally large, and more work is needed to determine feasibility of approach for lower accelerations

-Thrust and mass are completely decoupled from the analysis, which is limiting in terms of finding suitable hardware that can deliver the controls needed, as currently presented.
\end{itshape}

Thank you for the comment.
The conclusion has been expanded with some discussion of the drawbacks of this research and avenues of future work.
The modified conclusion is copied below.

\begin{correction}
In this paper, an optimal transfer process which combines concepts of reachability and \Poincare section is used to generate transfer between planar periodic orbits in the three-body problem.
The \Poincare section allows for trajectory design on a lower dimensional phase space and simplifies the process.
The indirect optimal control formulation enables straightforward method of incorporating additional path and control constraints.
However, the use of optimal control techniques leads to open loop trajectories that are not robust to model uncertainties or disturbances.
Furthermore, our approach relies heavily on the relative simplicity of the PCRTBP and this method is much more challenging for the non-planar case.

@@

ASSESSMENT:

The central premise of the paper - using optimization of Poincare distance as a guiding motivation to make low-thrust reachability analysis more tractable - is interesting, useful, and to my knowledge reasonably novel.  Furthermore, the quality of the writing is good (at local scales) and there is an appropriate degree of literature review.  However, there are two major problems with this paper.
\end{itshape}
\begin{itemize}

\item 
    \begin{itshape}
First, in terms of composition, large portions of the paper are spent on lengthy development of foundational work that is well-known within the astrodynamics community and available in textbooks. 
\end{itshape}

    Thank you for the comment, and we agree that there is significant repetition of material previously covered in other sources.
    The manuscript has been reduced and reorganized to enhance the clarity and clearly separate previously material and the contributions of this paper.
    The introduction has been reduced and the repetition and discuss on the variational integrator has been removed.
    The second section has been retitled to ``Problem Formulation and Mathematical Background'' and has subsections which cover all of the background material, namely the planar circular restricted three body problem, Jacobi integral, invariant manifolds and \Poincare maps, the variational integrator.
The next sections introduce the optimal control formulation for the reachability analysis and the numerical examples, followed by conclusions.
The section on variational integrators has been reduced and now only discusses the main developments of the integrator for the PCRTBP and omits the review of variational principles.

@@

\item 
    \begin{itshape}
        Secondly, the contributions and findings of the authors feel highly preliminary and do not seem to evaluate their method at significant depth.
    \end{itshape}

    Thank you for the comment. 
    The numerical examples are meant to demonstrate the use of this approach and the feasibility of using the reachability set on a \Poincare section to determine transfer trajectories.
    The examples give two demonstrations of finding a transfer trajectory in the PCRTBP and demonstrate combining several iterations of the reachability set for a transfer.

@@

\item
    \begin{itshape}
While there are the beginnings of a good article within this manuscript, it is not sufficiently close to publishable form to be accepted.  I recommend that the paper be rejected and that the authors push their investigation a bit further.  The eventual publication should focus less on background and should provide more insight into the fundamental properties of their method, using more rigorously constructed and analyzed examples. 
\end{itshape}

The authors believe that the given two nontrivial numerical examples illustrate the efficacy of the proposed approach in designing low-thrust maneuvers through reachability set analysis on the \Poincare section. 
Furthermore, the manuscript has been completely reorganized to emphasize the contributions and clearly separate the background material. 
In particular,~\cref{sec:simulation} includes additional paragraphs and figures for the geostationary to periodic orbit transfer.
These additional figures demonstrate the ability to chain together multiple iterations of the reachability set computation to generate a nontrivial transfer trajectory.
Combined together, the numerical examples demonstrate the ability to depart from the vicinity of the Earth to the vicinity of the Moon using our proposed method of coupling low-thrust maneuvers with reachability set analysis.

@@

\item 
    \begin{itshape}
I can sympathize that there was quite a long road to get to the point of understanding this topic and generating results in it. I hope that the authors will not find this verdict too discouraging, as I believe that stronger results are well within their reach.
\end{itshape}
\end{itemize}

SPECIFIC FEEDBACK:
\begin{enumerate}
    \item
    \begin{itshape}
*** Bloat
In the introduction, the exposition of the problem could be considerably condensed, given the venue of publication. Additionally, the somewhat sprawling discussion jumps back and forth, and many key points seem to be repeated more than necessary.  It needs to be organized into titled subsections with specific scope.
\end{itshape}
